---
title: "Homework 3"
author: "Jessica Flynn"
output: github_document
---

```{r setup}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1 

```{r load data}
data("instacart")
```

This data set contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. Observations are at the level of items in orders by users, , meaning there is 1 row for each product ordered in each order. There are user /  order variables `user_id`, `order_id`, `order_dow` and `order_hour_of_day`. There are also item variables -- `product_name`, `aisle`, `department`, and numeric codes for `product_id`, `aisle_id` and `department_id`. There are also some variables regarding whether someone re-ordered an item (`reordered`), what number order it is for a user (`order_number`) and how many days since their prior order (`days_since_prior_order`).

How many aisles are there, and which aisles are the most items ordered from?

There are a total for `r instacart %>% count(aisle) %>% nrow()` aisles. The top 3 aisles that the most items are ordered from are output in the table below.

```{r aisles_and_items}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n)) %>% 
  slice_head(n = 3) %>% 
  knitr::kable()
```


Plot for aisles with more than 10000+ items ordered. 

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate( 
    aisle = factor(aisle), 
    aisle = fct_reorder(aisle, n)
    ) %>%
  ggplot(aes(x = aisle, y = n)) + 
  labs(
    title = "Aisles with 10,000+ Items Ordered",
    x = "Aisle",
    y = "Number of Items Order") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  geom_point()
```

Below is a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits".

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```


Below is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour =  mean(order_hour_of_day, na.rm = TRUE)) %>% 
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour
  ) %>% 
  knitr::kable()
```

## Problem 2 

```{r, eval=FALSE}
accel_df = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>%
  pivot_longer() %>% 
  mutate(weekend = )
```

table of 35 days, total activity in each day. do week #1 and dow in each to make more readable. use factors to re-order days. check class of minutes

activity count y axis, minute on x axis, geom_line, color for dow,

## Problem 3

```{r}
library(p8105.datasets)
data("ny_noaa")
```
change tmin/tmax from 10ths of a degree to a degree

part 2 - group by and summarize + ggplot


part 3 - 2 different plot and merge them using patchwork