---
title: "Homework 3"
author: "Jessica Flynn"
output: github_document
---

```{r setup}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1 

```{r load data}
data("instacart")
```

This data set contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. Observations are at the level of items in orders by users. There are user /  order variables `user_id`, order ID, order dat and order hour. There are also item variables -- name, aisle, department, and some numeric codes. POLISH THIS TEXT

How many aisles are there, and which aisles are the most items ordered from?
```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Add a plot for aisles with more than 10000 items ordered. 

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate( 
    aisle = factor(aisle), 
    aisle = fct_reorder(aisle, n)
    ) %>%
  ggplot(aes(x = aisle, y = n)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  geom_point()
```

This is a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits"

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```


This is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour =  mean(order_hour_of_day, na.rm = TRUE)) %>% 
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour
  ) %>% 
  knitr::kable()
```

